---
title: "2020 Canadian employment income by group"
output: ioslides_presentation
date: "2024-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(haven)
library(dplyr)
library(forcats)
library(ggplot2)
```

## How does my salary compare? {data-background=#dff0e7}

The [2020 Canadian employment income by group](https://sagarana.shinyapps.io/income_percentile/) shiny app uses public microdata from Statistics Canada to:

-   show distribution of employment income by citizenship, gender, home ownership, and first language
-   calculate the percentile of an income entered by the user, showing differences across groups

## About the data

[2021 Census of Canada public microdata](https://hdl.handle.net/11272.1/AB2/1WTDOP) provides 2020 income data for a sample of nearly 1 million people in Canada. 

For meaningful salary comparisons the app uses the *employment income* variable, `EmpIn`.

For reproducibility, when uncommented this code will download, unzip, and load the data:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#download.file("https://abacus.library.ubc.ca/api/access/datafile/
#:persistentId?persistentId=hdl:11272.1/AB2/1WTDOP/HIYEIA",
#"CENSUS_2021.zip", mode="wb")
#unzip("CENSUS_2021.zip")

data2021 <- read_sav("cen_ind_2021_pumf_v2.sav")
```

## Preparing the data

The [2021 Census Individuals PUMF User Guide](https://hdl.handle.net/11272.1/AB2/1WTDOP/EV7NLM) documents each variable and informs data preparation.

Only a handful of variables are used in the app, and the data is limited to ages 18-69 (about 530K rows). Here's a sample table from the prepared dataset. 

```{r, message=FALSE, warning=FALSE, comment=NA}
census_subset_2021 <- data2021 %>%
        select(EmpIn, AGEGRP, TENUR, Citizen, FOL, Gender) %>%
        filter(AGEGRP > 6 & AGEGRP < 18) %>% #employment ages, keep respondents aged 18-69
        filter(EmpIn != 88888888 & EmpIn != 99999999) %>% #remove na
        as_factor()   #for better labels and data handling
        
levels(census_subset_2021$TENUR)[2] <- "Renter"
```

```{r, echo=TRUE}
table(census_subset_2021$Citizen, census_subset_2021$Gender)
```
<span style="color:grey; font-size:0.7em; margin-top:1em; font-style:italic"><b>Note:</b> the user guide explains that "individuals in the category 'non-binary persons' are distributed into the other two gender categories and are denoted by the '+' symbol."</span>

---- 

Here is app output for a reference salary of CAD $60,000 (dotted line) grouped by home ownership.

```{r}
mycols <- c("EmpIn", "TENUR")
max_disp_sal <- 150
salary <- 60
                
display_df <- census_subset_2021 %>%
         select(one_of(mycols)) %>%
         rename(Group = 2) # Change second column name to 'Group'
                
pctl <- NULL
for (i in levels(display_df$Group)) {
         df_filtered <- filter(display_df, Group == i)
         pctl <- c(pctl,round(ecdf(df_filtered$EmpIn)(salary*1000),2))
}
                
intersects <- data.frame(Group=levels(display_df$Group)) %>%
        mutate(input_salary=salary) %>%
        mutate(percentile=pctl)
                
ggplot(NULL, aes()) +
        geom_boxplot(data=display_df, aes(x=Group, y=EmpIn/1000, color=Group)) + 
        coord_cartesian(ylim = c(-70, max_disp_sal)) +
        geom_hline(yintercept=salary, linetype="dashed", color="slategrey") +
        geom_point(data=intersects, aes(x=Group, y=input_salary)) +
        geom_text(data=intersects, aes(x=Group, y=input_salary, label=percentile), hjust=-0.3, vjust=-0.5) +
        xlab("Group") +
        ylab("Employment income (thousands)") + 
        theme_light(base_size=9) +
        theme(axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.ticks.x=element_blank()) +
        scale_colour_brewer(palette = "Dark2")

```

In 2020 CAD $60,000 was the 80th percentile among renters.