---
title: 'Developing Data Products: Assignment 2'
output: ioslides_presentation
date: "2024-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The data (mtcars)

Graphs on the next slides use three variables from the `mtcars` data frame: mpg, hp, and cyl. Below are their values for the first 5 observations.

```{r, results='hide', warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(plotly)
```

```{r, echo=TRUE}
mtcars %>%
        select(c(mpg,hp,cyl)) %>%
        head(5)
```

## Initial graph (ggplot2)

The first step is to build a plot using the `ggplot2` package and to assign it to the variable `plot1`

```{r pressure}
plot1 <- mtcars %>%
        ggplot(aes(x=hp, y=mpg, color=as.factor(cyl))) +
        geom_point() +
        xlab("Horsepower") + 
        ylab("Miles/gallon") +
        labs(color = "Cylinders", title="MPG and HP by number of cylinders")
plot1
```

## Interactive graph (plotly)

Plotly's `ggplotly` function enhances `plot1` with interactivity (_hover over graph for options_)

```{r, warning=FALSE}
ggplotly(plot1) %>%
add_trace(
    type="scatter", 
    mode="markers",
    name="",
    x = mtcars$hp,
    y = mtcars$mpg,
    opacity = 0,
    text = rownames(mtcars),
    hovertemplate = paste('<b>%{text}</b><br>',
                          '<i>mpg:</i> %{y}<br>',
                        '<i>hp:</i> %{x}'),
    hoverlabel = list(bgcolor = "white"),
    showlegend = FALSE
  )

```